Mechanical Detection of Dipole-Dipole Interactions \\between Electronic Spins in a Solid

### 2020-11-01 Mes papiers/Méca CR/Meca_cross_SI.tex
tata
demenagement

@@

\begin{figure}[!ht]
  \centering \scalebox{0.32}{\includegraphics{CR_meca_22}}
  \caption{Mechanical detection of a degeneracy in a [100]$_\perp$ plane. a) Theoretical frequencies for the four $\ket{0} \to \ket{-1}$ transitions in dashed lines and mechanical ESR spectra vertically. b) Photoluminescence of the NV centers as a function of the scanning magnetic field (blue crosses) and gaussian fit (orange line). c) Simulated population in $\ket{m_s=0}$ for the stationary state with (plain) or without (dashed) taking into account the decrease of the T$_1$ induced by the cross-relaxations.
  d) Reflected signal off the diamond as a function of the scanning magnetic field (blue crosses) and double gaussian fit (orange line) e) Simulated torque applied by the spins on the diamond, with (plain) or without(dashed) taking into account the cross-relaxations.}
	\label{CR_22}
\end{figure}

@@

Our physical interpretation of the angular signal we observe, as well as the dispersive profile of the torque is that in this case, for symmetry reason, the magnetic torque generated by the four classes of NV is not modified when we are exactly at resonance, since all four classes are affected identically. What happens when you are near, but not at, resonance is that the classes will not be identically depolarized : looking at Fig. \ref{CR_22}a), we can see that the two classes of higher frequency are always slightly closer to each other than the two classes of lower frequency (we can see the slope of the frequencies being a bit smaller for the two upper classes), resulting in a bit more depolarization for these two classes, except when they are exactly at resonance. This interpretation accounts both for the overall shape of the responses, and the comparatively weakness of the phenomenon.

\section{Simulation details}
In this part we will discuss the method used to simulate the average torque as well as the $\ket{m_s=0}$ population of the stationary state. The numerical solving of the master equations was performed thanks to the Quantum Toolbox in Python (QuTiP) \citep{qutip1} \citep{qutip2}.

@@

As for the torque, we use a semi-classical formula :
\begin{equation*}
\mathbf \Gamma = N_0 \gamma_e <\hat{\mathbf S}> \times \mathbf B,
\end{equation*}
where $N_0 \approx 10^9$ is an estimate of the number of spins in our sample based on the average size and NV density of our diamonds, $\gamma_e$ is the gyromagnetic ratio of the electron and $<\hat{\mathbf S}> =\mathrm{Tr}(\rho_s \mathbf{\hat S})$ is the averaged spin vector in the stationary sate, averaged again on the four possible orientations of NV.
This formula assumes that the spin dynamics is faster than the dynamics of the motion of the diamond, which is the case in our experiments.

In our plots in Fig. X and Fig. 3 of the main ext, we only represent one spatial component (e.g. $\Gamma_x$) of the torque, but the three components behave similarly.
\bibliography{trap}

>>>>

\section{Principle of the mechanical detection}

\subsection{Torque sensing with a levitating diamond}

The way we can experimentally measure torques applied on a levitating diamond is by looking at the displacement of equilibrium caused by them. Here is a short model detailing it :

We model the trap as a pure harmonical potential, both for the center of mass and for the rotational degrees of freedom of the diamond (REF qq part), with trapping frequencies $\omega_T \approx 1$ kHz that we can experimentally measure. If we focus on a single rotational degree, we can write the torque generated by the trap as $\Gamma_T^\theta=-K(\theta-\theta_{eq})$, where $K=I \omega_t^2$ is the stiffness of the trap, $I$ being the moment of inertia of the diamond.

The application of an external torque $\Gamma_{\mathrm{ext}}$ on the diamond will therefore shift the angular position at equilibrium in such a way that : 
\begin{align*}
-K(\theta-\theta_{\mathrm{eq}}) + \Gamma_{\mathrm{ext}} &= -K(\theta-\theta_{\mathrm{eq}}') \\
\delta\theta = \theta_{\mathrm{eq}}'-\theta_{\mathrm{eq}} &= \frac{\Gamma_{\mathrm{ext}}}{K}=\frac{\Gamma_{\mathrm{ext}}}{I \omega_t^2}
\end{align*}

In our case, $\Gamma_{\mathrm{ext}}$ is the magnetic torque exerted by the NV$^-$ spins directly on the diamond, such that we can write $\Gamma_{\mathrm{ext}} = N_{NV} \langle \Gamma_{\mathrm{1 spin}} \rangle$  where $\langle \Gamma_{\mathrm{1 spin}} \rangle = \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B \approx 10^{-28}$ Nm is the classical magnetic torque applied by one spin (see Section  \ref{Simu}).

By using the inertia moment formula of a sphere : $I=\frac{2}{5}m r^2$, we can then rewrite the angular displacement as $$ \delta\theta = \frac{\langle\Gamma_{\mathrm{1 spin}}\rangle [NV^-]}{\frac{2}{5}m_Cr^2\omega_T^2} \approx 10^{-3} rad$$ where $[NV^-] \approx 5 \cdot 10^{-6}$ (5 ppm) is the number of NV centers per atoms in the crystal, $m_C \approx 2 \cdot 10^{-26}$ kg is the average weight of a carbon atom (we assume that the bulk of the diamond weight comes from carbon atoms), $r= 7.5$ $\mu$m is the typical radius of our diamonds and $\omega_T = 1$ kHz is the typical value of the trap frequency.

It should be noted that the main uncertainty comes here from the diamond size, which along the other uncertainties can change the expected result by an order of magnitude.

\subsection{Change of torque at the cross-relaxation}

@@

%Faire 2 lignes avec sur la première ligne l'ESR et le trajet sur la carte du champ mag
\begin{figure}[!ht]
  \centering \scalebox{0.32}{\includegraphics{CR_meca_22}}
  \caption{Mechanical detection of a degeneracy in a [100]$_\perp$ plane. a) Theoretical frequencies for the four $\ket{0} \to \ket{-1}$ transitions in dashed lines and mechanical ESR spectra vertically. b) Photoluminescence of the NV centers as a function of the scanning magnetic field (blue crosses) and gaussian fit (orange line). c) Simulated population in $\ket{m_s=0}$ for the stationary state with (plain) or without (dashed) taking into account the decrease of the T$_1$ induced by the cross-relaxations.
  d) Reflected signal off the diamond as a function of the scanning magnetic field (blue crosses) and double gaussian fit (orange line) e) Simulated torque applied by the spins on the diamond, with (plain) or without(dashed) taking into account the cross-relaxations.}
	\label{CR_22}
\end{figure}

@@

Our physical interpretation of the angular signal we observe, as well as the dispersive profile of the torque is that in this case, for symmetry reason, the magnetic torque generated by the four classes of NV is not modified when we are exactly at resonance, since all four classes are affected identically. What happens when being near, but not at, resonance is that the classes will not be identically depolarized : looking at Fig. \ref{CR_22}a), we can see that the two classes of higher frequency are always slightly closer to each other than the two classes of lower frequency (we can see the slope of the frequencies being a bit smaller for the two upper classes), resulting in a bit more depolarization for these two classes, except when they are exactly at resonance. This interpretation accounts both for the overall shape of the responses, and the comparatively weakness of the phenomenon.

\section{Simulation details}
\label{Simu}
In this part we will discuss the method used to simulate the average torque as well as the $\ket{m_s=0}$ population of the stationary state. The numerical solving of the master equations was performed thanks to the Quantum Toolbox in Python (QuTiP) \citep{qutip1} \citep{qutip2}.

@@

As for the torque, we use a semi-classical formula :
\begin{equation*}
\mathbf \Gamma = N_0 \gamma_e \langle\hat{\mathbf S}\rangle \times \mathbf B,
\end{equation*}
where $N_0 \approx 10^9$ is an estimate of the number of spins in our sample based on the average size and NV density of our diamonds, $\gamma_e$ is the gyromagnetic ratio of the electron and $\langle\hat{\mathbf S}\rangle =\mathrm{Tr}(\rho_s \mathbf{\hat S})$ is the averaged spin vector in the stationary sate, averaged again on the four possible orientations of NV.
This formula assumes that the spin dynamics is faster than the dynamics of the motion of the diamond, which is the case in our experiments.

In our plots in Fig. \ref{CR_22} and Fig. 3 of the main ext, we only represent one spatial component (e.g. $\Gamma_x$) of the torque, but the three components behave similarly.
\bibliography{trap}
